disaggregation: An \proglang{R} package for Bayesian spatial disaggregation modelling

### 2019-11-18 article.Rnw
Add latest MCMC run to paper. Delete unnecessary plots

The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package. Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm. It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}. The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took x hours. This number of iterations was chosen to be the minimum required for the value of Rhat to drop below 1.05 for all model parameters. The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds. Fitted parameter values for both of these methods are given in the table below.

@@

# Running the MCMC algorithm for 8 hours
start <- Sys.time()
mcmc_out <- tmbstan(m$obj, chains = 4, iter = 8000, warmup = 2000,
                     cores = getOption('mc.cores', 4))
end <- Sys.time()
print(end - start)

@@

%
\begin{table}
\centering
\begin{tabular}{l|ll|ll}
 & \multicolumn{2}{|c|}{MCMC (x hours - TBC)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
Parameter & Mean & SD & Mean & SD \\
\hline
Intercept & -3.06 & 0.42 & -3.12 & 0.27 \\
Slope 1 & 0.17 & 0.29 & 0.24 & 0.27 \\
Slope 2 & 0.31 & 0.22 & 0.30 & 0.21\\
Slope 3 & -0.34 & 0.22 & -0.37 & 0.20 \\
Slope 4 & -0.24 & 0.26 & -0.24 & 0.21 \\
iid $log(\tau)$ & 1.11 & 0.28 & 1.10 & 0.27 \\
$log(\sigma)$ & -3.35 & 0.60 & -3.55 & 0.59 \\
$log(\rho)$ & 0.74 & 0.34 & -0.56 & 0.30 \\
\end{tabular}
\end{table}
%

%
\begin{figure}[t!]
\centering
\includegraphics[width=\linewidth]{figures/mcmc_long.pdf}
\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for x hours.}
\label{fig:mcmclong}
\end{figure}
%

>>>>

The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package. Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm using four chains. It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}. The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 94 hours. This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations and doubling the number of iterations each time, until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters. The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds. Fitted parameter values for both of these methods are given in the table below.

@@

# Running the MCMC algorithm for 94 hours
start <- Sys.time()
mcmc_out <- tmbstan(m$obj, chains = 4, iter = 8000, warmup = 2000,
                     cores = getOption('mc.cores', 4))
end <- Sys.time()
print(end - start)

@@

%
\begin{table}
\centering
\begin{tabular}{l|ll|ll}
 & \multicolumn{2}{|c|}{MCMC (94 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
Parameter & Mean & SD & Mean & SD \\
\hline
Intercept & -3.09 & 0.33 & -3.12 & 0.27 \\
Slope 1 & 0.20 & 0.28 & 0.24 & 0.27 \\
Slope 2 & 0.32 & 0.21 & 0.30 & 0.21\\
Slope 3 & -0.36 & 0.20 & -0.37 & 0.20 \\
Slope 4 & -0.24 & 0.23 & -0.24 & 0.21 \\
iid $log(\tau)$ & 1.08 & 0.27 & 1.10 & 0.27 \\
$log(\sigma)$ & -3.34 & 0.59 & -3.55 & 0.59 \\
$log(\rho)$ & 0.72 & 0.31 & -0.56 & 0.30 \\
\end{tabular}
\end{table}
%

%
\begin{figure}[t!]
\centering
\includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 94 hours.}
\label{fig:mcmclong}
\end{figure}
%
